Active Learning for SAT Solver Benchmarking

### 2022-10-10 paper/paper.tex
Address comments

We transform the runtimes into discrete runtime classes which are determined on a per instance basis.
For each instance $e \in \mathcal{I}$, we use a single-link hierarchial clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of the $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
For instances with many timeouts, we have to manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows.
$$\gamma_k(a, e) = j ~\Leftrightarrow~ r(a, e) \in C_j$$

@@

\section{Evaluation}
\label{sec:eval}
Before evaluating our approach end-to-end as described in Section~\ref{sec:exdesign}, we discuss the choice of the runtime-discretization method and the machine-learning predictor.
Thereafter, we look at the results of our active-learning approach.
Selection decisions of the aforementioned approach also reveal the importance of different problem-instance families to our model.
Instance families comprise instances that are derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolio creation.
%JB: instance families not really motivated till now, only appear in abstract and in text regarding stratification
%TF: Added a sentence for clarification. We could introduce more details in section 4.3.

\subsection{Runtime Prediction}
\label{sec:evalprel}
%JB: if we dedicate a whole section to these results, we should probably rename this section (not "preliminary" in the typical sense); for TACAS, I guess this can be cut a bit
%TF: cut the section down a bit, renamed to "runtime prediction"
An exhaustive grid search of all active-learning approaches outlined in Section~\ref{sec:main} and different runtime-prediction models is infeasible.
Therefore, we use the runtime-prediction model that produced the best results for all further experiments.
We compare runtime predictors by looking at the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} of the classification task.
MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.

@@

\subsection{Experimental Results}
\subsubsection{Tuning our algorithm.}
Our main end-to-end experiments follow the evaluation framework that has been introduced in Section~\ref{sec:evalframe}.
Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches listed in Section~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
The plotted lines represent the Pareto front for each possible choice of ranking approaches (Fig.~\ref{fig:annitraincolorranking}), selection approaches (Fig.~\ref{fig:annitraincolorselection}), and stopping criteria (Fig.~\ref{fig:annitraincolorstopping}).
%JB: as far as I understand it, the lines only compare some top-level hyperparameters, but for each of these hyperparemeters, take the best (= Pareto front) out of some more fine-grained settings (the hyperparameters of the hyperparameters, so to speak), so it's no really "each possible choice" in the sense of actual, individual configurations, but rather a combination of configurations
Colors indicate the different hyper-parameter instantiations.
To be more specific, they represent the Pareto front of all configurations with the respective instantiations.

>>>>

We transform the runtimes into discrete runtime classes which are determined on a per instance basis.
For each instance $e \in \mathcal{I}$, we use a single-link hierarchial clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of the $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
For instances with many timeouts, we have to manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
%TF: maybe a bit unclear. We do not manually fix anything. There is always one cluster with only time-outs if there are any.
The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows.
$$\gamma_k(a, e) = j ~\Leftrightarrow~ r(a, e) \in C_j$$

@@

\section{Evaluation}
\label{sec:eval}

Before evaluating our approach end-to-end as described in Section~\ref{sec:exdesign}, we discuss the choice of the runtime-discretization method and the machine-learning predictor.
Thereafter, we look at the results of our active-learning approach.
Selection decisions of the aforementioned approach also reveal the importance of different problem-instance families to our model.
Instance families comprise instances that are derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolio creation.
%JB: instance families not really motivated till now, only appear in abstract and in text regarding stratification
%TF: Added a sentence for clarification. We could introduce more details in section 4.3.

\subsection{Runtime Prediction}
\label{sec:evalprel}

%JB: if we dedicate a whole section to these results, we should probably rename this section (not "preliminary" in the typical sense); for TACAS, I guess this can be cut a bit
%TF: cut the section down a bit, renamed to "runtime prediction"
An exhaustive grid search of all active-learning approaches outlined in Section~\ref{sec:main} and different runtime-prediction models is infeasible.
Therefore, we use the runtime-prediction model that produced the best results for all further experiments.
We compare runtime predictors by looking at the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} of the classification task.
MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.

@@

\subsection{Experimental Results}
\subsubsection{Tuning our algorithm.}

Our main end-to-end experiments follow the evaluation framework that has been introduced in Section~\ref{sec:evalframe}.
Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches listed in Section~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
The plotted lines represent the best performing configurations per ranking approach (Fig.~\ref{fig:annitraincolorranking}), per selection approach (Fig.~\ref{fig:annitraincolorselection}), and per stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
%JB: as far as I understand it, the lines only compare some top-level hyperparameters, but for each of these hyperparemeters, take the best (= Pareto front) out of some more fine-grained settings (the hyperparameters of the hyperparameters, so to speak), so it's no really "each possible choice" in the sense of actual, individual configurations, but rather a combination of configurations
Colors indicate the different hyper-parameter instantiations.
To be more specific, they represent the Pareto front of all configurations with the respective instantiations.
